---
title: "Network of Bigrams"
output:
  html_notebook:
    df_print: paged
---
* Modified by JJ
* Original one from prof. Rei
* Date: Nov 16, 2021


Reference book: 
<https://www.tidytextmining.com/>

```{r, message=FALSE, warning=FALSE}
library(tidytext)
library(tidyverse)
```

We will use text data built from the different entries in the News page of Florida Polytechnic University <https://floridapoly.edu/news/>

```{r, message=FALSE}
polynews <- read_csv("https://raw.githubusercontent.com/reisanar/datasets/master/poly_news_FL20.csv"); head(polynews)
```

```{r}
dim(polynews)
```

We may be interested in visualizing all of the relationships among words simultaneously, rather than just the top few at a time. As one common visualization, we can arrange the words into a network, or _graph_. Here we refer to a graph not in the sense of a visualization, but as a combination of connected nodes. A graph can be constructed from a tidy object since it has three variables

- `from`: the node an edge is coming from

- `to`: the node an edge is going towards

- `weight`: A numeric value associated with each edge


The `igraph` [package](https://igraph.org/) has many powerful functions for manipulating and analyzing networks. `igraph` is a collection of network analysis tools with the emphasis on efficiency, portability and ease of use. `igraph` is open source and free. 


One way to create an `igraph` object from tidy data is the `graph_from_data_frame()` function, which takes a data frame of edges with columns for `"from"`, `"to"`, and edge attributes (in this case `n`):

```{r, message=FALSE}
library(igraph)
```

First, perform tokenization and remove stop-words:

```{r}
poly_filtered <-  polynews %>%
  unnest_tokens(bigram, news_summary, token = "ngrams", n = 2) %>%
  separate(bigram, into = c("word1","word2"), sep = " ") %>%
  filter(!word1 %in% stop_words$word, !is.na(word1)) %>%
  filter(!word2 %in% stop_words$word, !is.na(word2))
  
  
```

We then find the most common combination of words:

```{r}
poly_counts <- poly_filtered %>% 
 count(word1, word2, sort = TRUE) %>% 
 filter(n > 5)

head(poly_counts)
```

Build graph from data frame:

```{r}
# filter for only relatively common combinations
bigram_graph <- poly_counts %>%
                filter(n > 1) %>%
                graph_from_data_frame()

bigram_graph
```

`igraph` has plotting functions built in, but they are not what the package is designed to do, so many other packages have developed visualization methods for graph objects. We recommend the `ggraph` [package](https://cran.r-project.org/web/packages/ggraph/index.html) (Pedersen 2017), because it implements these visualizations in terms of the _grammar of graphics_, which we are already familiar with from `ggplot2`.

```{r}
library(ggraph)
```

```{r}
set.seed(2017) # tackling the random ness

# set arrow
a <- grid::arrow(
      type = "closed",
      length = unit(.10,"inches"))

# plot graph
ggraph(bigram_graph, layout = "fr") +
    geom_edge_link(aes(edge_alpha = n),
    show.legend = FALSE, arrow = a,
    end_cap = circle(.07,'inches')) +
    geom_node_point(color = "#53316B",
    size = 4) +
    geom_node_text(aes(label = name),
    vjust = 1.5, hjust = 0.2)
    theme_void()
```

We conclude with a few polishing operations to make a better looking graph:

- We add the `edge_alpha` aesthetic to the link layer to make links transparent based on how common or rare the bigram is.

- We add directionality with an arrow, constructed using `grid::arrow()`, including an `end_cap` option that tells the arrow to end before touching the node.

- We tinker with the options to the node layer to make the nodes more attractive (larger, blue points)

- We add a theme that is useful for plotting networks, `theme_void()`

