---
title: "Simple Operations and Summaries"
subtitle: "Getting familiar with R"  
output: html_notebook
---

Let us try to reproduce one of the examples discussed during our lecture. Below we define two vectors, `x` and `y`. Notice the use of the `c()` function, used for combining values into a vector.

```{r}
# define x
x <- seq(1:10)
# define y
y <- x^2
```

If we want to find the mean and standard deviation of `x` and `y`, we can use the commands below:

```{r}
# mean of x
mean(x)
# mean of y
mean(y)
```


```{r}
# standard deviation of x
sd(x)
# standard deviation of y
sd(y)
```

Now let us check the value of correlation between `x` and `y`:

```{r}
cor(x,y)
```

(as discussed during the lecture, the (Pearson) correlation coefficient compute here, measures of the strength and direction of the _linear_ relationship between two variables - if any-)




## Datasets? 

R comes with a variety of built-in datasets that we can use to test different algorithms or simply practice our data analysis skills. 

One example is the `mtcars` dataset: the data extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973-74 models).

To check how the dataset looks like, simply type the name of the object (in this case `mtcars`) in a R _chunk of code_:

```{r}
mtcars
```


Typically we just want to see the first (or last) set of observations to get a quick sense of the type of information we are dealing with:

```{r}
# use the head() function to print the first 6 rows
head(mtcars)
```


```{r}
# use the tail() function to print the last 6 rows
tail(mtcars)
```


Formally, `mtcars` is called a **data frame**, where rows represents observations (records/samples) and columns represent features (variables/attributes)


### Summarizing a dataset

The `summary()` function is a powerful function that returns summary statistics for every variable in a given data frame:

```{r}
summary(mtcars)
```

