---
title: "Apriori algorithm: small example"
output: html_notebook
---

We will use the `arules` package to generate association rules and the `arulesViz` package to visualize them.

```{r, message=FALSE, warning=FALSE}
library(arules)
library(arulesViz)
library(tidyverse)
```

Documentation on the `arules` and `arulesViz` packages can be found here: 

<https://cran.r-project.org/web/packages/arules/vignettes/arules.pdf>

<https://cran.r-project.org/web/packages/arulesViz/vignettes/arulesViz.pdf>



## Example 1

In this example we will define the **transaction** from a _list_ object:

```{r}
# define lists
shopping_list <- list(
  c("flowers", "card", "soda"),
  c("toy", "flowers", "balloon", "candy"),
  c("card", "candy", "flowers"),
  c("toy", "balloons", "soda"),
  c("flowers", "card", "soda")
)
  
  
```

We now change the class of the list into a _transactions_ class object (the `transactions` class represents transaction data used for mining itemsets or rules):

```{r}
# coerce as transaction
shopping_tr <- as(shopping_list, "transactions")
```

The apriori algorithm is used below to generate rules from our shopping list:

```{r}
# uses the apriori() function
shopping_rules <- apriori(shopping_tr,
                          parameter = list(support = .1,
                                           confidence = .25,
                                           minlen = 2))
```

We use the `inspect()` function to check the top 3 rules based on _confidence_


```{r}
# inspect by confidence
inspect(head(sort(shopping_rules, by = "confidence"), 3))
image(shopping_tr)
```


## Example 2

In this example we create transactions from a data frame:


```{r}
## creating transactions from data.frame
students_df <- data.frame(
  age = factor(c(6, 8, NA, 9, 16)),
  grade = factor(c("A", "C", "F", NA, "C")),
  pass = c(T, T, F, T, T)
)
```

Again, we _coerce_ the data frame as a `transactions` object:

```{r}
## coerce
students_tr <- as(students_df, "transactions")
```

We can inspect this new object:

```{r}
# inspect
inspect(students_tr)
```

Below the apriori algorithm is used with its default values:


```{r}
students_rules <- apriori(students_tr)
```

Let us inspect the generated rules this time based on the top 6 rules by _lift_:

```{r}
students_rules %>%
  sort( . , by = "lift") %>%
  head() %>%
  inspect()
```


## Example 3

First, let us create transactions from a list:

```{r}
publix <- list(
  c("milk", "tea", "cake"),
  c("eggs", "tea", "cold drink"),
  c("milk", "eggs", "tea", "cold drink"),
  c("eggs", "cold drink"),
  c("juice")
)
```

Convert list to transactions class:

```{r}
publix_tr <- as(publix, "transactions")
```

An item frequency plot can be created using the `itemFrequencyPlot()` function:

```{r}
itemFrequencyPlot(publix_tr)
```


Use apriori algorithm with support set to 0.4:

```{r}
publix_rules <- apriori(publix_tr, parameter = list(support = .4))
```



Inspect the rules:

```{r}
inspect(publix_rules)
```

Let us visualize these rules using tools from the `arulesViz` package.



```{r, message=FALSE, warning=FALSE}
plot(publix_rules, method = "graph", control = list(type = "items"), engine = "htmlwidget")
```








