---
title: "Apriori algorithm: small example"
output: html_notebook
---

We will use the `arules` package to generate association rules and the `arulesViz` package to visualize them.

```{r, message=FALSE, warning=FALSE}
library(arules)
library(arulesViz)
library(tidyverse)
```

Documentation on the `arules` and `arulesViz` packages can be found here: 

<https://cran.r-project.org/web/packages/arules/vignettes/arules.pdf>

<https://cran.r-project.org/web/packages/arulesViz/vignettes/arulesViz.pdf>



## Example 1

In this example we will define the **transaction** from a _list_ object:

```{r}
# define lists
shopping_list <- 
  
  
```

We now change the class of the list into a _transactions_ class object (the `transactions` class represents transaction data used for mining itemsets or rules):

```{r}
# coerce as transaction

```

The apriori algorithm is used below to generate rules from our shopping list:

```{r}
# uses the apriori() function

  
```

We use the `inspect()` function to check the top 3 rules based on _confidence_


```{r}
# inspect by confidence

```

```{r}

```

```{r}

```


## Example 2

In this example we create transactions from a data frame:


```{r}
## creating transactions from data.frame

```

Again, we _coerce_ the data frame as a `transactions` object:

```{r}
## coerce

```

We can inspect this new object:

```{r}
# inspect

```

Below the apriori algorithm is used with its default values:


```{r}

```

Let us inspect the generated rules this time based on the top 6 rules by _lift_:

```{r}

```


## Example 3

First, let us create transactions from a list:

```{r}

```

Convert list to transactions class:

```{r}

```

An item frequency plot can be created using the `itemFrequencyPlot()` function:

```{r}


```


Use apriori algorithm with support set to 0.4:

```{r}


```



Inspect the rules:

```{r}

```

Let us visualize these rules using tools from the `arulesViz` package.



```{r, message=FALSE, warning=FALSE}

```








