---
title: "SVD illustration"
output: html_notebook
---

Below we write a _function_ that let us create a [Hilbert matrix](https://en.wikipedia.org/wiki/Hilbert_matrix) of any dimension.

```{r}
# create Hilbert matrix of size n 
hilbert <- function(n) { 
  i <- 1:n
  1 / outer(i - 1, i, "+") 
  }
```

Let us consider the Hilbert matrix of size $n=9$. We will create such matrix, and then extract the first 6 columns of it:

```{r}
# create Hilbert matrix of size 9
hil_nine <- hilbert(9)
# select all rows, and columns 1 through 6
X <- hil_nine[ , 1:6]
```


Below we print the entire matrix $X$ with $m=9$ rows and $n=6$ columns:

```{r}
# print matrix X
print(X)
```

One way to calculate the **singular value decomposition** (SVD) of the matrix $X$ is to use the `svd()` function in R:

```{r}
# compute SVD using the svd() function
s <- svd(X)
```


Notice that `s` is an object of class _list_, containing 3 main pieces of information:

- `d`, a vector containing the singular values of matrix `X`

- `u`, a matrix whose columns contain the _left singular vectors_ of `X`

- `v`, a matrix whose columns contain the _right singular vectors_ of `X`

```{r}
# find out the class of the svd object
class(s)
```

Mathematically, the singular value decomposition (SVD) of a matrix $X$ looks like: 

$$
X = U \Sigma V^T
$$

where $\Sigma = \text{diag}(d)$ is a diagonal matrix containing the **singular values** of matrix $X$. Both matrices, $U$ and $V$ are **orthogonal**, that is, $UU^T$ and $VV^T$ are equal to the _identity matrix_ (of the associated dimensions)




We now extract the components from the SVD object:

```{r}
# matrix of left singular vectors
myU <- 
# diagonal matrix with singular values
mySigma <- 
# matrix of right singular vectors
myV <- 
```

We can confirm this decomposition (also sometimes called factorization), using the matrix multiplication operator in R, `%*%`, and the transpose operator `t()`

```{r}
## U * Sigma * V^T

```


Check the magnitude of the singular values:

```{r}
# print singular values of X

```


Suppose we focus only on the first 3 singular values, then using the first 3 columns for both $U$ and $V$

```{r}
# reduced U
red_U <- 
red_V <- 
red_S <- 
```

an approximation to $X$ is created: 

```{r}
# approximation to matrix X
approx_matrix <- 
# print matrix
approx_matrix
```

Notice that by using just 3 singular values, we obtain a good approximation of the original matrix $X$ (this verifies the use of SVD for compression purposes)

```{r}
# find the difference between X and 
# its approximation with 3 singular values

```


