---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
fuel <- read_csv("https://raw.githubusercontent.com/reisanar/datasets/master/fuel.csv",
                 col_types = cols())
```


We can also use the `dplyr::sumarize()` function to get some summaries for certain variables. Try to generate a summary that shows the minimum and maximum values for the variables `year`, `annual_fuel_cost_ft1` and `annual_consumption_in_barrels_ft1`:


> This is the best way I can think of using summarise for this. The code is a pain to write and a bit difficult to read. The output is also difficult to read.

```{r}
fuel %>%
  summarize(yearMin = min(year), yearMax = max(year),
            fuelCostMin = min(annual_fuel_cost_ft1), fuelCostMax = max(annual_fuel_cost_ft1),
            consMin = min(annual_consumption_in_barrels_ft1), consMax = max(annual_consumption_in_barrels_ft1))
```


> Of course, it seems like the easiest thing to do would to create a vector or list of the columns. Creating a list doesn't work at all and using a vector just grabs the min/max total value of all three columns.

```{r}
fuel %>%
  summarize(min(c(year, annual_fuel_cost_ft1, annual_consumption_in_barrels_ft1)),
            max(c(year, annual_fuel_cost_ft1, annual_consumption_in_barrels_ft1)))
```


> However, using `summary()` instead of `summarise()` gives you a neatly formatted list of summary statistics

```{r}
# insert code here
fuel %>%
  select(year, annual_fuel_cost_ft1, annual_consumption_in_barrels_ft1) %>%
  summary()
```


> If you only want the `min` and `max` values, it's a bit more work, but it can still be done more simply than with `summarize()`

```{r}
summ <- fuel %>%
  select(year, annual_fuel_cost_ft1, annual_consumption_in_barrels_ft1) %>%
  summary()
summ[c(1,6),]
```


> Or an awful one-liner without any tidyverse

```{r}
summary(fuel[, c("year", "annual_fuel_cost_ft1", "annual_consumption_in_barrels_ft1")])[c(1,6),]
```


> I guess I'm just really wondering what the best way to use `summarise()` would be and what the advantages of doing it that way would be over just using base R