---
title: "Homework Assignment 2"
subtitle: "Dimensionality Reduction using PCA"
author: "Gus Lipkin - glipkin6737@floridapoly.edu"
output: html_notebook
---


Load the `tidyverse` and `factoextra` packages for this assignment.

```{r, message = FALSE, warning=FALSE, comment=FALSE}
library(tidyverse)
library(factoextra)
```


# Problem 1 (25 points)

In this problem you are asked to find a recent research article that uses **Principal Component Analysis (PCA)** (or a variation of it).

Share:

- The name of the paper, authors, and affiliations.

- A short description of the main results (check the abstract and conclusions of the paper)

- A direct link to the work.

You could use the University's library for this, or simply use Google Scholar for a list of the many papers that have used this methodology in recent years.




# Problem 2 (25 points)

Consider the `USArrests` dataset, which contains statistics, in arrests per 100,000 residents for assault, murder, and rape in each of the 50 US states in 1973. Also given is the percent of the population living in urban areas.

```{r}
head(USArrests)
```

(a) Perform PCA on the four numerical variables in this dataset.

```{r}
arrests_pca <- prcomp(USArrests, scale. = T)
```


(b) Create a bi-plot (in the PC1-PC2 coordinate system) and explain your results. According to your findings: Do Florida, Nevada, California, and Michigan have anything in common? How about New Hampshire, Maine, and North Dakota?

```{r}
fviz_pca_biplot(arrests_pca, 
                select.ind = list(names = c("Florida", "Nevada", "California",
                                            "Michigan", "New Hampshire", "Maine",
                                            "North Dakota")))
```
> Comments: Maine and North Dakota have the most in common and are close to New Hampshire, while the other four states are similar in some respects but still quite different.


# Problem 3 (25 points)

In this problem you will use data from the [EA Sports FIFA 18](https://en.wikipedia.org/wiki/FIFA_18) videogame. 

- Get the dataset `fifa18.csv`: 

```{r, message=FALSE}
fifa18 <- read_csv("https://raw.githubusercontent.com/reisanar/datasets/master/fifa18.csv", col_types = cols())
```

- Below is a sample of observations in this dataset:

```{r}
head(fifa18, 8)
```

(a) Use the numerical attributes in this dataset to perform PCA. Get a summary of your PCA results (use the `summary()` function as shown in examples). Comment on your results. 

```{r}
fifa18 <- fifa18[!duplicated(fifa18$name),]
# dataframe with only numerical attributes
myfifa18 <- select(fifa18, -c(1:3))
tibble::remove_rownames(myfifa18)
rownames(myfifa18) <- fifa18$name
```

> Extra comments: I had to engage in shenanigans here so that I could get player names to show up on the visualization. I just got rid of any rows where a player appeared a second or more time and then set the row names to the player names.

Perform PCA: 

```{r}
# code here - include as many chunks of code as needed
fifa_pca <- prcomp(myfifa18, scale. = T)
summary(fifa_pca)
```

> Comments: It's interesting how the amount of explanation by principal component tapers off. The first one is over 50%, then to get to 75% you need either 3 or 4 (depends if you're willing to round). To get to 90%, you need to go all the way out to PC9 and then to PC16 for 95%.



(b) Produce at least one **data visualization** to explain the results on PCA. (You can use the `factoextra` package for this and follow the examples discussed in class)

```{r}
# code here - include as many chunks of code as needed
fviz_pca_biplot(fifa_pca, 
                select.ind = list(name = c("Cristiano Ronaldo", "L. Messi", "Neymar", "Gus Outomouro", "Gus Ledes")))
```

> Comments: This was fun. The three names I know, Ronaldo, Messi, and Neymar, are all grouped together where I guess all the good players go. The two other players I've never heard of before that are named Gus share some similar properties, but not in such great quantities as the three well-known players.




# Problem 4 (25 points)

(a) Read the Toyota Corollas dataset on sales during late summer of 2004 in the Netherlands. It has 1436 records containing details on 38 attributes, including Price, Age, Kilometers, HP, and other specifications. 

```{r, message=FALSE}
toyota <- read_csv("https://raw.githubusercontent.com/reisanar/datasets/master/ToyotaCorolla.csv", col_types = cols())
```


- Print the first 6 rows of the dataset.

```{r}
head(toyota)
```


- Identify (to the best of your knowledge) the categorical variables.

> Comments: `model`, `fuel_type`, and `color`. I'm not sure if indicator variables count, but if they do, `mfr_guarantee` and variables to the right.


(b) Create a new data frame with **some of the numerical attributes** in this dataset (ignoring the categorical variables, and think about which variables are probably categorical variables even when represented by a number). 

Perform **Principal Component Analysis (PCA)**, and **comment on your results**.

```{r}
# code here - include as many chunks of code as needed
# you could use the example above to select the collection of variables you want
myToyota <- toyota[,c("price", "age_08_04", "km", "hp", "quarterly_tax", "weight", "cc")]
toyota_pca <- prcomp(myToyota, scale. = T)
summary(toyota_pca)
```

> Comments: 


(c) Produce at least one **data visualization** to explain the results on PCA. (You can use the `factoextra` package for this and follow the examples discussed in class)

```{r}
# code here - include as many chunks of code as needed

```


## Problem 5 (20 points)

> **Required for graduate students**. _Extra-credit (+20 points) for undergraduate students_.

You will use data generated by the Spotify API on songs nominated to the _Grammy Award for Song of the Year_.

Find the data in this [GitHub repo](https://github.com/reisanar/datasets/blob/master/grammy_softy_20.csv), and read it into RStudio using:

```{r, message=FALSE}
# read grammy data
soty <- read_csv("https://raw.githubusercontent.com/reisanar/datasets/master/grammy_softy_20.csv", col_types = cols())
```

(a) Perform PCA on this dataset, using all the **numerical attributes** available. 

```{r}

```


(b) Summarize your findings and comment on your results. 

```{r}

```


(c) Apply PCA again to a smaller number of numerical attributes (briefly explain your reasoning for the selection of the numerical attributes you consider for this part of the problem)

```{r}

```


(d) Create at least one data visualization of your PCA results.

```{r}

```

